---
title: "Relative Humidity Analysis"
output: html_notebook
---

# 1. Imports

```{r, include=FALSE}
library(corrplot)
library(tidyverse)
library(sf)
Sys.setlocale("LC_TIME", "en_US")
```

## 1.1 Readings datasets

Let's start by importing the processed relative humidity readings:

```{r}
humid_readings <- read.csv(
    "processed_dataset/Umidita_relativa_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)
```

Let's apply some small preprocessing to the imported dataset:

```{r}
humid_readings <- humid_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )
```

## 1.2 Sensors/stations dataset

We also need the processed Lombardy stations for the analysis:

```{r}
sensors <- read.csv(
    "processed_dataset//Stazioni_Idro_Nivo_Meteorologiche.csv",
    header = TRUE,
    strip.white = TRUE
)
```

Let's extract the necessary sensors:

```{r}
humid_sensors <- sensors |>
    filter(Tipologia == "Umidità Relativa") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Provincia = as.factor(Provincia)
    )
```


# 2. Time-Coverage Analysis

Let's see the time coverage of data by each type of sensor. First of, let's extract the month from each reading:

```{r}
humid_readings <- humid_readings |>
    mutate(Month = month(Data, label=TRUE, abbr=FALSE))
```

We also need to join the readings dataset with the sensors dataset:

```{r}
humid_sensor_readings <- left_join(
    humid_readings,
    humid_sensors,
    join_by(IdSensore == IdSensore)
)
```

Finally, let's compute the time-coverage graph:

```{r}
humid_sensor_readings |> 
    filter(Stato == "VA") |>
    ggplot(
        aes(x = Data, y = IdSensore)
    ) +
    geom_point(
        aes(color = Month)
    ) +
    facet_wrap(
        . ~ Provincia,
        scales = "free_y"
    ) +
    scale_x_date(
        date_breaks = "1 month",
        labels = scales::label_date_short()
    ) +
    labs(
        x = element_blank(),
        y = element_blank()
    ) + theme_minimal() +
    theme(
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", linetype = "dashed")
    )
```

Data is available for the entire year. Some sensors have missing days or entire months, probably either due to malfunctions that required on-site repairs, or because they were commissioned during the time period.

# 3. Relative Humidity Evolution

## 3.1 Correlation

Let's try to understand what could be related to the evolution of the relative humidity, by drawing a correlation plot. We first need to compute the correlation matrix on the desired features:

```{r}
humid_sensor_readings_tmp <- humid_sensor_readings |>
    filter(Stato == "VA") |>
    mutate(Month = month(Data)) |>
    select(
        Media,
        Month,
        Quota,
        lat,
        lng
    )

humid_cor = cor(humid_sensor_readings_tmp, use = "complete.obs")
```

Finally, we just need to plot the correlation graph:

```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(humid_cor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45,
         col = col(200), addCoef.col = "black", cl.pos = "n", order = "AOE")
```

Doesn't look like there is any strong relationship between humidity and altitude at which the reading is made. Let's see it within a scatterplot:

```{r}
humid_sensor_readings |>
    filter(Stato == "VA") |>
    ggplot(
        aes(x = Quota, y = Media)
    ) +
    geom_point(
        aes(color = Media)
    ) +
    stat_smooth(
        method = "lm"
    ) +
    labs(
        x = "Average Relative Humidity (%)",
        y = "Altitude (m)"
    ) +
    scale_colour_distiller(palette = "Greens") +
    theme_minimal()
```

```{r}
humid_sensor_readings |>
    filter(
        Stato == "VA",
        !(Month %in% c("June", "July", "August", "September"))
    ) |>
    ggplot(
        aes(x = Quota, y = Media)
    ) +
    geom_point(
        aes(color = Media)
    ) +
    geom_smooth(
        method = "lm", se = FALSE,
        color = "gray20", linetype = "dashed"
    ) +
    facet_wrap(
        . ~ Month,
        ncol = 3
    ) +
    labs(
        x = "Average Relative Humidity (%)",
        y = "Altitude (m)"
    ) +
    theme_minimal() +
    scale_colour_distiller(palette = "Greens") +
    theme(
        legend.position = "inside",
        legend.justification = c(1,0), legend.position.inside = c(1,0),
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
```

## 3.1 Annual relative humidity trend

Compute mean readings for each province:

```{r}
monthly_humid <- humid_sensor_readings |>
    mutate(
        Month = month(Data),
        Year = year(Data)
    ) |>
    group_by(Provincia, NomeProvincia, Month, Year) |>
    summarise(
        Media = if (all(is.na(Media))) NA else round(mean(Media, na.rm = TRUE)),
        Max = if (all(is.na(Max))) NA else max(Max, na.rm = TRUE),
        Min = if (all(is.na(Min))) NA else min(Min, na.rm = TRUE),
        .groups = "drop"
    ) |>
    mutate(Stato = ifelse(is.na(Media), "NA", "VA"))
```

Now we just have to plot the monthly trend for each province:

```{r}
gg_humid_monthly <- monthly_humid |>
    filter(Stato == "VA") |>
    mutate(
        Data = make_date(Year, Month, day = 1),
        NomeProvincia = paste(NomeProvincia, " (", Provincia, ")", sep = "")
    ) |>
    ggplot(
        aes(x = Data, y = Media)
    ) +
    geom_point(
        aes(color = NomeProvincia),
        size = 1
    ) +
    geom_line(
        aes(color = NomeProvincia)
    ) +
    geom_smooth(
        method = "lm", se = FALSE,
        color = "gray20", linetype = "dashed"
    ) +
    facet_wrap(
        . ~ NomeProvincia,
        ncol = 3
    ) +
    scale_x_date(
        date_breaks = "1 month",
        labels = scales::label_date_short()
    ) +
    labs(
        x = element_blank(),
        y = "Humidity (%)",
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
gg_humid_monthly
```

```{r, include=FALSE}
ggsave(
    filename = "images/humid_monthly.jpg",
    plot = gg_humid_monthly,
    width = 24,
    height = 14,
    dpi = 300,
    units = "cm"
)
```

In May, it looks like there was a drastic decrease in average solar irradiance across almost the entire region. There were indeed multiple "extreme" meteorological events during that month.

## 3.2 Altitude relative humidity trend

Let's try to plot a boxplot of solar irradiance over altitude:

```{r}
humid_sensor_readings$AltitudeRange <- cut(
    humid_sensor_readings$Quota,
    breaks = seq(0, 3200, by = 800),
    labels = c("0-800m", "800-1600m", "1600-2400m", "2400-3200m")
)

humid_altitude_periods <- humid_sensor_readings |>
    filter(
        Stato == "VA",
        AltitudeRange != "0-800m",
        !(Month %in% c("July", "August", "September"))
    ) |>
    mutate(
        Month = factor(Month, levels = c(
            "October", "November", "December", 
             "January", "February", "March", 
             "April", "May", "June"
        ))
    ) |>
    ggplot(
        aes(x = AltitudeRange, y = Media)
    ) +
    geom_boxplot(
        aes(fill = AltitudeRange)
    ) +
    facet_wrap(
        . ~ Month,
        ncol = 3
    ) +
    labs(
        x = element_blank(), y = "Humidity (%)"
    ) +
    scale_fill_brewer(palette = "Greens") +
    theme_minimal() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
humid_altitude_periods
```

```{r, include=FALSE}
ggsave(
    filename = "images/humid_altitude_periods.jpg",
    plot = humid_altitude_periods,
    width = 24,
    height = 14,
    dpi = 300,
    units = "cm"
)
```

Let's try to compute the monthly average for each sensor:

```{r}
monthly_humid_sensors <- humid_sensor_readings |>
    mutate(
        Month = month(Data),
        Year = year(Data)
    ) |>
    group_by(IdSensore, NomeStazione, Quota, Month, Year) |>
    summarise(
        Media = if (all(is.na(Media))) NA else round(mean(Media, na.rm = TRUE)),
        Max = if (all(is.na(Max))) NA else max(Max, na.rm = TRUE),
        Min = if (all(is.na(Min))) NA else min(Min, na.rm = TRUE),
        .groups = "drop"
    ) |>
    mutate(Stato = ifelse(is.na(Media), "NA", "VA"))
```

Finally, let's plot the time-series at different altitude ranges:

```{r}
monthly_humid_sensors$AltitudeRange <- cut(
    monthly_humid_sensors$Quota,
    breaks = seq(0, 3200, by = 800),
    labels = c("0-800m", "800-1600m", "1600-2400m", "2400-3200m")
)

monthly_humid_sensors |>
    filter(Stato == "VA") |>
    mutate(
        Data = make_date(Year, Month, day = 1)
    ) |>
    ggplot(
        aes(x = Data, y = Media)
    ) +
    geom_density(
        stat = "identity",
        aes(color = IdSensore, fill = IdSensore),
        alpha = 0.15
    ) +
    facet_wrap(
        . ~ AltitudeRange,
        ncol = 3
    ) +
    scale_x_date(
        date_breaks = "1 month",
        labels = scales::label_date_short()
    ) +
    labs(
        x = element_blank(),
        y = "Average Humidity (%)",
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
```

Relative Humidity and altitude are independent (as was already proven by the correlation matrix).