---
title: "Station Readings"
output: html_notebook
---

# 1. Imports

```{r, include=FALSE}
library(corrplot)
library(geosphere)
library(tidyverse)
library(sf)
Sys.setlocale("LC_TIME", "en_US")
```

## 1.1 Readings datasets

Let's start by importing all processed readings:

```{r}
snow_readings <- read.csv(
    "processed_dataset/Altezza_neve_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)

irrad_readings <- read.csv(
    "processed_dataset/Radiazione_Globale_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)

temp_readings <- read.csv(
    "processed_dataset/Temperatura_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)

humid_readings <- read.csv(
    "processed_dataset/Umidita_relativa_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)

prcp_readings <- read.csv(
    "processed_dataset/Precipitazioni_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)
```

Let's apply some small preprocessing to the imported datasets:

```{r}
snow_readings <- snow_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )

irrad_readings <- irrad_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )

temp_readings <- temp_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )

humid_readings <- humid_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )

prcp_readings <- prcp_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )
```

## 1.2 Sensors/stations dataset

We also need the processed Lombardy stations for the analysis:

```{r}
sensors <- read.csv(
    "processed_dataset//Stazioni_Idro_Nivo_Meteorologiche.csv",
    header = TRUE,
    strip.white = TRUE
)
```

Let's also extract the different types of sensors:

```{r}
snow_sensors <- sensors |>
    filter(Tipologia == "Altezza Neve") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        IdStazione = as.factor(IdStazione),
        Provincia = as.factor(Provincia),
        NomeProvincia = as.factor(NomeProvincia)
    )

irrad_sensors <- sensors |>
    filter(Tipologia == "Radiazione Globale") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        IdStazione = as.factor(IdStazione),
        Provincia = as.factor(Provincia),
        NomeProvincia = as.factor(NomeProvincia)
    )

temp_sensors <- sensors |>
    filter(Tipologia == "Temperatura") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        IdStazione = as.factor(IdStazione),
        Provincia = as.factor(Provincia),
        NomeProvincia = as.factor(NomeProvincia)
    )

humid_sensors <- sensors |>
    filter(Tipologia == "Umidità Relativa") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        IdStazione = as.factor(IdStazione),
        Provincia = as.factor(Provincia),
        NomeProvincia = as.factor(NomeProvincia)
    )

prcp_sensors <- sensors |>
    filter(Tipologia == "Precipitazione") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        IdStazione = as.factor(IdStazione),
        Provincia = as.factor(Provincia),
        NomeProvincia = as.factor(NomeProvincia)
    )
```

## 1.3 GeoJSON map

Import the Lombardy GeoJSON map:

```{r}
lombardy <- st_read("geojson/lombardy_.geojson")
```


# 2. Combining Everything

## 2.1 Merging all types of readings by station

At this point, we need to combine each readings with the corresponding sensor information. This can be done with a left join on each reading dataset by the sensor/station dataset:

```{r}
snow_readings <- snow_readings |> left_join(snow_sensors, by = "IdSensore")
irrad_readings <- irrad_readings |> left_join(irrad_sensors, by = "IdSensore")
temp_readings <- temp_readings |> left_join(temp_sensors, by = "IdSensore")
humid_readings <- humid_readings |> left_join(humid_sensors, by = "IdSensore")
prcp_readings <- prcp_readings |> left_join(prcp_sensors, by = "IdSensore")
```

We also need to extract only the relevant information for the analysis, and rename a few columns for clarity:

```{r}
snow_data <- snow_readings |> select(
    IdStazione,
    Data,
    AvgSnowHeight = Media,
    MinSnowHeight = Min,
    MaxSnowHeight = Max,
    NomeStazione,
    Quota,
    Provincia,
    NomeProvincia,
    lng,
    lat
)

irrad_data <- irrad_readings |> select(
    IdStazione,
    Data,
    AvgSolarIrradiance = Media,
    MinSolarIrradiance = Min,
    MaxSolarIrradiance = Max
)

temp_data <- temp_readings |> select(
    IdStazione,
    Data,
    AvgTemperature = Media,
    MinTemperature = Min,
    MaxTemperature = Max
)

humid_data <- humid_readings |> select(
    IdStazione,
    Data,
    AvgHumidity = Media,
    MinHumidity = Min,
    MaxHumidity = Max
)

prcp_data <- prcp_readings |> select(
    IdStazione,
    Data,
    TotalPrecipitation = Valore
)
```

Finally, let's merge all datasets based on the station IDs:

```{r}
station_data <- snow_data |>
    inner_join(irrad_data, by = c("IdStazione", "Data")) |>
    inner_join(temp_data, by = c("IdStazione", "Data")) |>
    inner_join(humid_data, by = c("IdStazione", "Data")) |>
    inner_join(prcp_data, by = c("IdStazione", "Data"))

station_data$IdStazione <- droplevels(station_data$IdStazione)
```

## 2.2 Deriving readings by surrounding stations

We're not in luck (porcoschifo), since only 21 out of 30 eligible stations (the ones equipped with a snow height sensor) have all other types of sensors available.

```{r}
complete_sta_ids <- unique(station_data$IdStazione)

complete_irrad_sta_ids <- snow_data |>
    inner_join(irrad_data, by = c("IdStazione", "Data")) |>
    pull(IdStazione) |>
    unique() |>
    droplevels()
missing_irrad_sta_ids <- factor(setdiff(unique(snow_data$IdStazione), complete_irrad_sta_ids))

complete_humid_sta_ids <- snow_data |>
    inner_join(humid_data, by = c("IdStazione", "Data")) |>
    pull(IdStazione) |>
    unique() |>
    droplevels()

missing_humid_sta_ids <- factor(setdiff(unique(snow_data$IdStazione), complete_humid_sta_ids))
```

Let's see it visually:

```{r}
sta_unequipped <- ggplot() + 
    geom_sf(
        data = lombardy,
        fill = "#c4e3ed",
        color = "black"
    ) + 
    geom_point(
        data = snow_sensors |> filter(IdStazione %in% complete_sta_ids),
        mapping = aes(x = lng, y = lat, color = "Fully-Equipped"),
        size = 2.5,
    ) + 
    geom_point(
        data = snow_sensors |> filter(IdStazione %in% missing_irrad_sta_ids),
        mapping = aes(x = lng, y = lat, color = "Missing Radiometer"),
        size = 2.5,
    ) + 
    geom_point(
        data = snow_sensors |> filter(IdStazione %in% missing_humid_sta_ids),
        mapping = aes(x = lng, y = lat, color = "Missing Hygrometer"),
        size = 2.5,
    ) +
    scale_color_manual(
        values = c(
            "Fully-Equipped" = "#04BF33",
            "Missing Radiometer" = "#F26D6D",
            "Missing Hygrometer" = "#5E9FF2"
        )
    ) +
    labs(
        x = element_blank(),
        y = element_blank(),
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        axis.text = element_blank(),
        legend.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
sta_unequipped
```

```{r, include=FALSE}
ggsave(
    filename = "images/stations_unequipped.png",
    plot = sta_unequipped,
    width = 24,
    height = 16,
    dpi = 300,
    units = "cm"
)
```

Luckily, only solar irradiance (8) and relative humidity (1) sensors are missing from each of these incomplete stations.

### 2.2.1 Deriving solar irradiance readings

Let's see the stations with missing solar irradiance sensors, and the surrounding stations equipped with solar irradiance sensors:

```{r}
complete_sta_ids <- snow_data |>
    inner_join(irrad_data, by = c("IdStazione", "Data")) |>
    pull(IdStazione) |>
    unique() |>
    droplevels()

missing_sta_ids <- factor(setdiff(unique(snow_data$IdStazione), complete_sta_ids))
```

```{r}
sta_irrad_incomplete <- ggplot() + 
    geom_sf(
        data = lombardy,
        fill = "#c4e3ed",
        color = "black"
    ) + 
    geom_point(
        data = irrad_sensors,
        mapping = aes(x = lng, y = lat, color = "Radiometer-Equipped"),
        size = 2.5,
    ) + 
    geom_point(
        data = snow_sensors |> filter(IdStazione %in% missing_sta_ids),
        mapping = aes(x = lng, y = lat, color = "Missing Radiometer"),
        size = 2.5,
    ) +
    scale_color_manual(
        values = c(
            "Radiometer-Equipped" = "#04BF33",
            "Missing Radiometer" = "#F26D6D"
        )
    ) +
    labs(
        x = element_blank(),
        y = element_blank(),
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        legend.text = element_text(size = 18),
        axis.text = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
sta_irrad_incomplete
```

```{r, include=FALSE}
ggsave(
    filename = "images/stations_irrad_incomplete.png",
    plot = sta_irrad_incomplete,
    width = 18,
    height = 18,
    dpi = 300,
    units = "cm"
)
```

Wouldn't it be nice if we could make an approximation of solar irradiance based on the surrounding stations? Let's first define a function to compute the distance between two geographic locations, based on their coordinates:

```{r}
# Function to calculate distance between two points (Haversine formula)
calculate_distance <- function(lat1, lon1, lat2, lon2) {
  distHaversine(c(lon1, lat1), c(lon2, lat2)) / 1000  # Convert to kilometers
}
```

Then compute these distances from the incomplete stations, and filter those within 15Km:

```{r}
# Extract stations
incomplete_sta <- snow_sensors |>
    filter(IdStazione %in% missing_sta_ids)

equipped_sta <- irrad_sensors

# Create a dataframe to store the distances
distance_df <- expand.grid(incomplete_sta$IdStazione, equipped_sta$IdStazione)
colnames(distance_df) <- c("IncompleteStation", "EquippedStation")

# Calculate distances
distance_df <- distance_df |>
    rowwise() |>
        mutate(
            Distance = calculate_distance(
              incomplete_sta$lat[IncompleteStation == incomplete_sta$IdStazione],
              incomplete_sta$lng[IncompleteStation == incomplete_sta$IdStazione],
              equipped_sta$lat[EquippedStation == equipped_sta$IdStazione],
              equipped_sta$lng[EquippedStation == equipped_sta$IdStazione]
            )
    ) |>
    filter(Distance <= 15)

# Extract surrounding stations
equipped_sta <- irrad_sensors |>
    filter(IdStazione %in% unique(distance_df$EquippedStation))
```

Let's visualize the surrounding equipped stations:

```{r}
sta_irrad_around <- ggplot() + 
    geom_sf(
        data = lombardy,
        fill = "#c4e3ed",
        color = "black"
    ) + 
    geom_point(
        data = equipped_sta,
        mapping = aes(x = lng, y = lat, color = "Radiometer In-Range"),
        size = 2.5,
    ) + 
    geom_point(
        data = incomplete_sta,
        mapping = aes(x = lng, y = lat, color = "Missing Radiometer"),
        size = 2.5,
    ) +
    scale_color_manual(
        values = c(
            "Radiometer In-Range" = "#D9B533",
            "Missing Radiometer" = "#F26D6D"
        )
    ) +
    labs(
        x = element_blank(),
        y = element_blank(),
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        legend.text = element_text(size = 18),
        axis.text = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
sta_irrad_around
```

```{r, include=FALSE}
ggsave(
    filename = "images/stations_irrad_around.png",
    plot = sta_irrad_around,
    width = 18,
    height = 18,
    dpi = 300,
    units = "cm"
)
```

```{r}
ggsave(
    filename = "images/stations_irrad_around.png",
    plot = sta_irrad_incomplete + sta_irrad_around,
    width = 36,
    height = 18,
    dpi = 300,
    units = "cm"
)
```

Now that we have the surrounding stations, we can compute an approximation of the solar irradiance for each incomplete station:

```{r}
approx_irrad_data <- distance_df |>
    inner_join(
        irrad_data,
        by = c("EquippedStation" = "IdStazione"),
        relationship = "many-to-many"
    ) |>
    group_by(IncompleteStation, Data) |>
    summarise(
        AvgSolarIrradiance = if (all(is.na(AvgSolarIrradiance))) -999 else round(mean(AvgSolarIrradiance, na.rm = TRUE)),
        MinSolarIrradiance = if (all(is.na(MinSolarIrradiance))) -999 else min(MinSolarIrradiance, na.rm = TRUE),
        MaxSolarIrradiance = if (all(is.na(MaxSolarIrradiance))) -999 else max(MaxSolarIrradiance, na.rm = TRUE),
        .groups = "drop"
    )

colnames(approx_irrad_data) <- c("IdStazione", "Data", "AvgSolarIrradiance", "MinSolarIrradiance", "MaxSolarIrradiance")
```

Then we merge all datasets based on the station IDs:

```{r}
incomplete_station_data <- snow_data |>
    filter(IdStazione %in% missing_sta_ids) |>
    inner_join(approx_irrad_data, by = c("IdStazione", "Data")) |>
    inner_join(temp_data, by = c("IdStazione", "Data")) |>
    inner_join(humid_data, by = c("IdStazione", "Data")) |>
    inner_join(prcp_data, by = c("IdStazione", "Data"))

incomplete_station_data$IdStazione <- droplevels(incomplete_station_data$IdStazione)
```

And finally add all the not-anymore-incomplete stations to the stations data:

```{r}
station_data <- bind_rows(station_data, incomplete_station_data)
station_data <- station_data |>
    distinct(IdStazione, Data, .keep_all = TRUE)
```

### 2.2.2 Deriving humidity readings

Let's see the stations with missing relative humidity sensors, and the surrounding stations equipped with relative humidity sensors:

```{r}
complete_sta_ids <- snow_data |>
    inner_join(humid_data, by = c("IdStazione", "Data")) |>
    pull(IdStazione) |>
    unique() |>
    droplevels()

missing_sta_ids <- factor(setdiff(unique(snow_data$IdStazione), complete_sta_ids))
```

```{r}
ggplot() + 
    geom_sf(
        data = lombardy,
        fill = "#c4e3ed",
        color = "black"
    ) + 
    geom_point(
        data = humid_sensors,
        mapping = aes(x = lng, y = lat, color = "Hygrometer-Equipped"),
        size = 2.5,
    ) + 
    geom_point(
        data = snow_sensors |> filter(IdStazione %in% missing_sta_ids),
        mapping = aes(x = lng, y = lat, color = "Missing Hygrometer"),
        size = 2.5,
    ) +
    scale_color_manual(
        values = c(
            "Hygrometer-Equipped" = "#04BF33",
            "Missing Hygrometer" = "#5E9FF2"
        )
    ) +
    labs(
        x = element_blank(),
        y = element_blank(),
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        axis.text = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
```

Compute the distances from the incomplete stations, and filter those within 15Km:

```{r}
# Extract stations
incomplete_sta <- snow_sensors |>
    filter(IdStazione %in% missing_sta_ids)

equipped_sta <- humid_sensors

# Create a dataframe to store the distances
distance_df <- expand.grid(incomplete_sta$IdStazione, equipped_sta$IdStazione)
colnames(distance_df) <- c("IncompleteStation", "EquippedStation")

# Calculate distances
distance_df <- distance_df |>
    rowwise() |>
        mutate(
            Distance = calculate_distance(
              incomplete_sta$lat[IncompleteStation == incomplete_sta$IdStazione],
              incomplete_sta$lng[IncompleteStation == incomplete_sta$IdStazione],
              equipped_sta$lat[EquippedStation == equipped_sta$IdStazione],
              equipped_sta$lng[EquippedStation == equipped_sta$IdStazione]
            )
    ) |>
    filter(Distance <= 15)

# Extract surrounding stations
equipped_sta <- humid_sensors |>
    filter(IdStazione %in% unique(distance_df$EquippedStation))
```

Let's visualize the surrounding equipped stations:

```{r}
ggplot() + 
    geom_sf(
        data = lombardy,
        fill = "#c4e3ed",
        color = "black"
    ) + 
    geom_point(
        data = equipped_sta,
        mapping = aes(x = lng, y = lat, color = "Hygrometer In-Range"),
        size = 2.5,
    ) + 
    geom_point(
        data = incomplete_sta,
        mapping = aes(x = lng, y = lat, color = "Missing Hygrometer"),
        size = 2.5,
    ) +
    scale_color_manual(
        values = c(
            "Hygrometer In-Range" = "#D9B533",
            "Missing Hygrometer" = "#5E9FF2"
        )
    ) +
    labs(
        x = element_blank(),
        y = element_blank(),
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        axis.text = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
```

Now that we have the surrounding stations, we can compute an approximation of the solar irradiance for each incomplete station:

```{r}
approx_humid_data <- distance_df |>
    inner_join(
        humid_data,
        by = c("EquippedStation" = "IdStazione"),
        relationship = "many-to-many"
    ) |>
    group_by(IncompleteStation, Data) |>
    summarise(
        AvgHumidity = if (all(is.na(AvgHumidity))) -999 else round(mean(AvgHumidity, na.rm = TRUE)),
        MinHumidity = if (all(is.na(MinHumidity))) -999 else min(MinHumidity, na.rm = TRUE),
        MaxHumidity = if (all(is.na(MaxHumidity))) -999 else max(MaxHumidity, na.rm = TRUE),
        .groups = "drop"
    )

colnames(approx_humid_data) <- c("IdStazione", "Data", "AvgHumidity", "MinHumidity", "MaxHumidity")
```

Then we merge all datasets based on the station IDs:

```{r}
incomplete_station_data <- snow_data |>
    filter(IdStazione %in% missing_sta_ids) |>
    inner_join(approx_humid_data, by = c("IdStazione", "Data")) |>
    inner_join(irrad_data, by = c("IdStazione", "Data")) |>
    inner_join(temp_data, by = c("IdStazione", "Data")) |>
    inner_join(prcp_data, by = c("IdStazione", "Data"))

incomplete_station_data$IdStazione <- droplevels(incomplete_station_data$IdStazione)
```

And finally add all the not-anymore-incomplete stations to the stations data:

```{r}
station_data <- bind_rows(station_data, incomplete_station_data)
station_data <- station_data |>
    distinct(IdStazione, Data, .keep_all = TRUE)
```

## 2.3 Saving to CSV

For saving in a CSV file, we would prefer to replace all `NA` values with `-999`:

```{r}
station_data <- station_data |>
    mutate(across(
        c(
            AvgSnowHeight,
            MinSnowHeight,
            MaxSnowHeight,
            AvgSolarIrradiance,
            MinSolarIrradiance,
            MaxSolarIrradiance,
            AvgTemperature,
            MinTemperature,
            MaxTemperature,
            AvgHumidity,
            MinHumidity,
            MaxHumidity,
            TotalPrecipitation
        ),
        ~ replace_na(., -999)
    ))
```

Not really necessary, but let's reorder the columns and rename them in Italian:

```{r}
final_data <- station_data |>
    select(
        IdStazione,
        Data,
        AvgSnowHeight,
        MinSnowHeight,
        MaxSnowHeight,
        AvgSolarIrradiance,
        MinSolarIrradiance,
        MaxSolarIrradiance,
        AvgTemperature,
        MinTemperature,
        MaxTemperature,
        AvgHumidity,
        MinHumidity,
        MaxHumidity,
        TotalPrecipitation
    ) |>
    rename(
        MediaAltezzaNeve = AvgSnowHeight,
        MinAltezzaNeve = MinSnowHeight,
        MaxAltezzaNeve = MaxSnowHeight,
        MediaIrraggiamento = AvgSolarIrradiance,
        MinIrraggiamento = MinSolarIrradiance,
        MaxIrraggiamento = MaxSolarIrradiance,
        MediaTemperatura = AvgTemperature,
        MinTemperatura = MinTemperature,
        MaxTemperatura = MaxTemperature,
        MediaUmidità = AvgHumidity,
        MinUmidità = MinHumidity,
        MaxUmidità = MaxHumidity,
        TotPrecipitazione = TotalPrecipitation
    ) |>
    arrange(IdStazione, Data)
```

Finally, save the computed station daily average readings in a separate .csv file:

```{r}
write.csv(
    final_data,
    file = "processed_dataset/Letture_Stazioni_dal_01062023_al_01062024.csv",
    sep = ",",
    quote = FALSE,
    row.names = FALSE
)
```