---
title: "Temperature Analysis"
output: html_notebook
---

# 1. Imports

```{r, include=FALSE}
library(corrplot)
library(tidyverse)
library(sf)
Sys.setlocale("LC_TIME", "en_US")
```

## 1.1 Readings datasets

Let's start by importing the processed temperature readings:

```{r}
temp_readings <- read.csv(
    "processed_dataset/Temperatura_dal_01062023_al_01062024.csv",
    header = TRUE,
    strip.white = TRUE,
    na.strings = c("-999")
)
```

Let's apply some small preprocessing to the imported dataset:

```{r}
temp_readings <- temp_readings |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Data = ymd(Data),
        Stato = as.factor(Stato)
    )
```

## 1.2 Sensors/stations dataset

We also need the processed Lombardy stations for the analysis:

```{r}
sensors <- read.csv(
    "processed_dataset//Stazioni_Idro_Nivo_Meteorologiche.csv",
    header = TRUE,
    strip.white = TRUE
)
```

Let's extract the necessary sensors:

```{r}
temp_sensors <- sensors |>
    filter(Tipologia == "Temperatura") |>
    mutate(
        IdSensore = as.factor(IdSensore),
        Provincia = as.factor(Provincia)
    )
```


# 2. Time-Coverage Analysis

Let's see the time coverage of data by each type of sensor. First of, let's extract the month from each reading:

```{r}
temp_readings <- temp_readings |>
    mutate(Month = month(Data, label=TRUE, abbr=FALSE))
```

We also need to join the readings dataset with the sensors dataset:

```{r}
temp_sensor_readings <- left_join(
    temp_readings,
    temp_sensors,
    join_by(IdSensore == IdSensore)
)
```

Finally, let's compute the time-coverage graph:

```{r}
temp_sensor_readings |> 
    filter(Stato == "VA") |>
    ggplot(
        aes(x = Data, y = IdSensore)
    ) +
    geom_point(
        aes(color = Month)
    ) +
    facet_wrap(
        . ~ Provincia,
        scales = "free_y"
    ) +
    scale_x_date(
        date_breaks = "1 month",
        labels = scales::label_date_short()
    ) +
    labs(
        x = element_blank(),
        y = element_blank()
    ) + theme_minimal() +
    theme(
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", linetype = "dashed")
    )
```

Data is available for the entire year. Some sensors have missing days or entire months, probably either due to malfunctions that required on-site repairs, or because they were commissioned during the time period.

# 3. Temperature Evolution

## 3.1 Correlation

Let's try to understand what could be related to the evolution of the snow height, by drawing a correlation plot. We first need to compute the correlation matrix on the desired features:

```{r}
temp_sensor_readings_tmp <- temp_sensor_readings |>
    filter(Stato == "VA") |>
    mutate(Month = month(Data)) |>
    select(
        Media,
        Month,
        Quota,
        lat,
        lng
    )

temp_cor = cor(temp_sensor_readings_tmp, use = "complete.obs")
```

Finally, we just need to plot the correlation graph:

```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(temp_cor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45,
         col = col(200), addCoef.col = "black", cl.pos = "n", order = "AOE")
```

Looks like there might be an inverse correlation between temperature and altitude at which the reading is made. Let's see it within a scatterplot:

```{r}
temp_sensor_readings |>
    filter(Stato == "VA") |>
    ggplot(
        aes(x = Quota, y = Media)
    ) +
    geom_point(
        aes(color = Media)
    ) +
    geom_smooth(
        method = "lm"
    ) +
    labs(
        x = "Average Temperature (°C)",
        y = "Altitude (m)"
    ) +
    scale_colour_distiller(palette = "Reds") +
    theme_minimal()
```

```{r}
temp_sensor_readings |>
    filter(
        Stato == "VA",
        !(Month %in% c("June", "July", "August", "September"))
    ) |>
    ggplot(
        aes(x = Quota, y = Media)
    ) +
    geom_point(
        aes(color = Media)
    ) +
    geom_smooth(
        method = "lm", se = FALSE,
        color = "gray20", linetype = "dashed"
    ) +
    facet_wrap(
        . ~ Month,
        ncol = 3
    ) +
    labs(
        x = "Average Temperature (°C)",
        y = "Altitude (m)"
    ) +
    theme_minimal() +
    scale_colour_distiller(palette = "Reds") +
    theme(
        legend.position = "inside",
        legend.justification = c(1,0), legend.position.inside = c(1,0),
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
```

## 3.1 Annual temperature trend

Compute mean readings for each province:

```{r}
monthly_temp <- temp_sensor_readings |>
    mutate(
        Month = month(Data),
        Year = year(Data)
    ) |>
    group_by(Provincia, NomeProvincia, Month, Year) |>
    summarise(
        Media = if (all(is.na(Media))) NA else round(mean(Media, na.rm = TRUE)),
        Max = if (all(is.na(Max))) NA else max(Max, na.rm = TRUE),
        Min = if (all(is.na(Min))) NA else min(Min, na.rm = TRUE),
        .groups = "drop"
    ) |>
    mutate(Stato = ifelse(is.na(Media), "NA", "VA"))
```

Now we just have to plot the monthly trend for each province:

```{r}
temp_monthly <- monthly_temp |>
    filter(Stato == "VA") |>
    mutate(
        Data = make_date(Year, Month, day = 1),
        NomeProvincia = paste(NomeProvincia, " (", Provincia, ")", sep = "")
    ) |>
    ggplot(
        aes(x = Data, y = Media)
    ) +
    geom_point(
        aes(color = NomeProvincia),
        size = 1
    ) +
    geom_line(
        aes(color = NomeProvincia)
    ) +
    geom_smooth(
        method = "lm", se = FALSE,
        color = "gray20", linetype = "dashed"
    ) +
    facet_wrap(
        . ~ NomeProvincia,
        ncol = 3
    ) +
    scale_x_date(
        date_breaks = "1 month",
        labels = scales::label_date_short()
    ) +
    labs(
        x = element_blank(),
        y = "Temperature (°C)",
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
temp_monthly
```

```{r, include=FALSE}
ggsave(
    filename = "images/temp_monthly.jpg",
    plot = temp_monthly,
    width = 24,
    height = 14,
    dpi = 300,
    units = "cm"
)
```

Looks like 2024 enters the summer season with much cooler air temperature, compared to 2023.

## 3.2 Altitude temperature trend

Let's try to plot a boxplot of temperature over altitude:

```{r}
temp_sensor_readings$AltitudeRange <- cut(
    temp_sensor_readings$Quota,
    breaks = seq(0, 3200, by = 800),
    labels = c("0-800m", "800-1600m", "1600-2400m", "2400-3200m")
)

temp_altitude_periods <- temp_sensor_readings |>
    filter(
        Stato == "VA",
        AltitudeRange != "0-800m",
        between(Media, -20, 40),
        !(Month %in% c("July", "August", "September"))
    ) |>
    mutate(
        Month = factor(Month, levels = c(
            "October", "November", "December", 
             "January", "February", "March", 
             "April", "May", "June"
        ))
    ) |>
    ggplot(
        aes(x = AltitudeRange, y = Media)
    ) +
    geom_hline(
        yintercept = 0, 
        linetype = "dashed", 
        color = "#F26D6D"
    ) +
    geom_boxplot(
        aes(fill = AltitudeRange)
    ) +
    facet_wrap(
        . ~ Month,
        ncol = 3
    ) +
    labs(
        x = element_blank(), y = "Temperature (°C)"
    ) +
    scale_fill_brewer(palette = "Reds") +
    theme_minimal() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
temp_altitude_periods
```

```{r, include=FALSE}
ggsave(
    filename = "images/temp_altitude_periods.jpg",
    plot = temp_altitude_periods,
    width = 24,
    height = 14,
    dpi = 300,
    units = "cm"
)
```

A strong relationship can be seen between temperature and altitude (as anyone could expect). Higher altitudes tend to show cooler temperatures, compared to lower altitudes.

Let's try to compute the monthly average for each sensor:

```{r}
monthly_temp <- temp_sensor_readings |>
    mutate(
        Month = month(Data),
        Year = year(Data)
    ) |>
    group_by(IdSensore, NomeStazione, Quota, Month, Year) |>
    summarise(
        Media = if (all(is.na(Media))) NA else round(mean(Media, na.rm = TRUE)),
        Max = if (all(is.na(Max))) NA else max(Max, na.rm = TRUE),
        Min = if (all(is.na(Min))) NA else min(Min, na.rm = TRUE),
        .groups = "drop"
    ) |>
    mutate(Stato = ifelse(is.na(Media), "NA", "VA"))
```

Finally, let's plot the time-series at different altitude ranges:

```{r}
monthly_temp$AltitudeRange <- cut(
    monthly_temp$Quota,
    breaks = seq(0, 3200, by = 800),
    labels = c("0-800m", "800-1600m", "1600-2400m", "2400-3200m")
)

temp_altitude_sensors <- monthly_temp |>
    filter(
        Stato == "VA",
        AltitudeRange != "0-800m",
        between(Media, -20, 20),
   ) |>
    mutate(
        Data = make_date(Year, Month, day = 1)
    ) |>
    ggplot(
        aes(x = Data, y = Media)
    ) +
    geom_density(
        stat = "identity",
        aes(color = IdSensore, fill = IdSensore),
        alpha = 0.15
    ) +
    facet_wrap(
        . ~ AltitudeRange,
        ncol = 3
    ) +
    scale_x_date(
        date_breaks = "1 month",
        labels = scales::label_date_short()
    ) +
    labs(
        x = element_blank(),
        y = "Average Temperature (°C)",
        color = element_blank()
    ) +
    theme_minimal() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 8),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "gray90", color = "gray90", size = 1)
    )
temp_altitude_sensors
```

```{r, include=FALSE}
ggsave(
    filename = "images/temp_altitude_sensors.jpg",
    plot = temp_altitude_sensors,
    width = 24,
    height = 8,
    dpi = 300,
    units = "cm"
)
```